A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-06-08 main.tex
Section 1 finished.
Fix typos.

\blue{181 words. The abstract should briefly summarize the contents of the paper in 150--250 words.}
\blue{コードとかはボールドを使ったほうが良いかもしれない。}
\red{Suspicious senetences are colored red.}
%
%
%
\red{Multistage Programming と Multistage Calculus をどう使い分けるか?}
\section{Introduction}

@@

Multistage programming enable us to generate and run fragments of code at runtime.
MetaOCaml\cite{oleg2014} provide features for multistage programming, which are quote, unquote, and running code.
Quote, written \verb|.< e >.| in MetaOCaml, makes a code value from \verb|e|.
Run, written \verb| run e | in MetaOCaml, runs a code value of \verb|e| and restore \verb|e|.
Unqote, written \verb| .~ e | in MetaOCaml, expands a code value of \verb|e|.
Unlike run, unquote is supposed to use in quote, that is, unquote cannot be used as alternative of run.
For example, the following MetaOCaml expression

@@

Cross-stage Persistence (CSP) is another primitive of multistage programming, which enable us to embed computed values into a code value.
This is an example of CSP.
\begin{verbatim}
        let plusone = fun x -> x + 1 in
        let a = plusone 2 in  .< a >.
\end{verbatim}
This program evaluates to \verb|.< 3 >.| not  \verb|.< plusone 2 >.|.
This is because the variable \verb|a| is introduced into a code fragment using CSP.
Therefore, the variable \verb|a| is embeded after \verb|a| is calculated.
CSP is very important when we write practival programs 
because CSP enables us to use library functions in code fragments.

@@

Multistage programming can also optimize vector calculation.
For example, \verb|add| function, which takes two vectors and return the sum of them,
is impremented with loop in many cases.
We can unroll \verb|add| function for a given vector length and optimize it with multistage programming.
\red{MetaOCamlで書いた長さ付きベクトルの例?}

@@

Although unrolled \verb|add| function is optimzied, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shoudn't use it for vectors of 3 lengths.
Otherwise, we will get Segmentation Fault error.
This problem is serious but existing type systems for multistage calculi cannot prevent it.

@@

As above, functions generated with multistage programming can take more restricted values as arguments.
When you optimize \verb|add| function for the length of 5, you should it only for 5 length vectors.
We introduce dependent types into a multistage calculus
so that the type sysytem can guarantee optimzied functions are used properly.

\subsubsection{Organization of the Paper}

The organization of this paper is following.

\section{Informal Overview of \LMD}

>>>>

\blue{181 words. The abstract should briefly summarize the contents of the paper in 150--250 words.}
\blue{コードとかはボールドを使ったほうが良いかもしれない。}
\red{Suspicious senetences are colored red.}
%
%
%
\red{Multistage Programming と Multistage Calculus をどう使い分けるか?}

\section{Introduction}

@@

Multistage programming enables us to generate and run fragments of code at runtime.
MetaOCaml\cite{oleg2014} provide features for multistage programming, which are quote, unquote, and running code.
Quote, written \verb|.< e >.| in MetaOCaml, makes a code value from \verb|e|.
Run, written \verb| run e | in MetaOCaml, runs a code value of \verb|e| and restore \verb|e|.
Unquote, written \verb| .~ e | in MetaOCaml, expands a code value of \verb|e|.
Unlike run, unquote is supposed to use in a quoted fragment of code, that is, unquote cannot be used as an alternative of run.
For example, the following MetaOCaml expression

@@

Cross-stage Persistence (CSP) is another primitive of multistage programming, which enable us to embed computed values into a code value.
This is an example of CSP.
\begin{verbatim}
        let plusone = fun x -> x + 1 in
        let a = plusone 2 in  .< a >.
\end{verbatim}
This program evaluates to \verb|.< 3 >.| not  \verb|.< plusone 2 >.|.
This is because the variable \verb|a| is introduced into a code fragment using CSP.
Therefore, the variable \verb|a| is embedded after \verb|a| is calculated.
CSP is very important when we write practical programs 
because CSP enables us to use library functions in code fragments.

@@

Multistage programming can also optimize vector calculation.
For example, \verb|add| function, which takes two vectors and return the sum of them,
is implemented with a loop in many cases.
We can unroll \verb|add| function for a given vector length and optimize it with multistage programming.
\red{MetaOCamlで書いた長さ付きベクトルの例?}

@@

Although unrolled \verb|add| function is optimized, we cannot use it for different vector length.
For example, when you optimize for the length of 5, you shouldn't use it for vectors of 3 lengths.
Otherwise, we will get a Segmentation Fault error.
This problem is serious but existing type systems for multistage calculi cannot prevent it.

@@

As above, functions generated with multistage programming can take more restricted values as arguments.
When you optimize \verb|add| function for the length of 5, you should it only for 5 length vectors.
We introduce dependent types into a multistage calculus
so that the type system can guarantee optimized functions are used properly.

\subsection{Organization of the Paper}

The organization of this paper is the following.
Section 2 gives an informal overview of \LMD.
Section 3 defines the syntax, type system, full reduction, evaluation contexts of \LMD.
Section 4 shows the properties of \LMD including Unique Decomposition.
Section 5 discusses future works.
Finally, Section 6 discusses related works.
\section{Informal Overview of \LMD}
